EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-30 draft.tex
Spell check. I think it's ready to go

Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously.
The fitness of each individual can be evaluated independently of 
others, and similarly, each population could evolve in isolation. Since 
earlier works, researchers have been proposing some form of parallelization
\cite{muhlenbein1988evolution} to increase the scalability of these algorithms.
The island model was one of the first techniques proposed for parallelization,
which lead to an increased performance \cite{gorges1990explicit,grosso1985computer}. 
The concept was to divide a large population into communicated subpopulations. 
Since then, other population-based algorithms have adopted the concept, 
and researchers have found additional benefits
besides the execution speed; these include avoiding a premature convergence and
maintaining the diversity of the global population \cite{li2015multi}. Researchers use the term
multi-population based methods when generally referring to techniques using
many populations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
% Comments below - Mario
Moreover, parallel implementations can run in two ways: synchronous and
asynchronous. When operations run in parallel but must maintain synchronism with
the other operations,  all operations need to finish to continue to the next
step. For instance, in a  parent/worker model, the parent needs to wait for all
workers to finish their operations before moving to the next iteration.  In
contrast, in an asynchronous execution,  operations are not synchronized with
each other. Following the previous example, now the parent operation can
continue to the next iteration, even if a worker has not finished its work.  In
the literature, we can find many asynchronous algorithms
\cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time,
and scalability. In the particular case of asynchronous and cloud-native
multi-population algorithms, recent works propose an asynchronous communication
through a central repository \cite{sofea:cec2012, JSON} or message queues
\cite{salza2019speed, guervos2018introducing}, in this work we follow this
practice.

@@

The EvoSwarm application uses configuration files to run the noiseless BOBB testbed
experiments.  As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, where the number of worker containers is
specified. Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The { \em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm. Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function will be tested. {\em Instances}  has a defualt value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of function evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

@@

As a conclusion, EvoSwarm is a framework designed and tested according
to industry standars which is able to bring many benefits to the
bioinspired algorithm community, by showing the possibility of
creating new, high performance, mixed algorithms, which can also be
easily deployed to the cloud, or locally to systems that simply need
to have Docker and Docker Compose installed.

It also opens new possibilities. From the dessign point of view, it
would be interesting to try and use automatic scaling, instead of
setting the number of workers by hand. This would need a certain
amount of redesign, the most important of which would be using
container orchestration systems such as Kubernetes or Docker
Swarm. Instead of setting a fixed number of workers, only the maximum
amount would need to be established, which could be done directly or
by setting an evaluation budget.

>>>>

Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously.
The fitness of each individual can be evaluated independently of 
others, and similarly, each population could evolve in isolation. Since 
earlier works, researchers have been proposing some form of parallelization
\cite{muhlenbein1988evolution} to increase the scalability of these algorithms.
The island model was one of the first techniques proposed for parallelization,
which lead to an increased performance \cite{gorges1990explicit,grosso1985computer}. 
The concept was to divide a large population into communicated subpopulations. 
Since then, other population-based algorithms have adopted the concept, 
and researchers have found additional benefits
besides the execution speed; these include avoiding a premature convergence and
maintaining the diversity of the global population \cite{li2015multi}. Researchers use the term
multi-population based methods when generally referring to techniques using
many populations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
% Comments below - Mario
Moreover, parallel implementations can run in two ways: synchronous and
asynchronous. When operations run in parallel but must maintain synchrony with
the other operations,  all operations need to finish to continue to the next
step. For instance, in a  parent/worker model, the parent needs to wait for all
workers to finish their operations before moving to the next iteration.  In
contrast, in an asynchronous execution,  operations are not synchronized with
each other. Following the previous example, now the parent operation can
continue to the next iteration, even if a worker has not finished its work.  In
the literature, we can find many asynchronous algorithms
\cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time,
and scalability. In the particular case of asynchronous and cloud-native
multi-population algorithms, recent works propose an asynchronous communication
through a central repository \cite{sofea:cec2012, JSON} or message queues
\cite{salza2019speed, guervos2018introducing}, in this work we follow this
practice.

@@

The EvoSwarm application uses configuration files to run the noiseless BBOB testbed
experiments.  As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, where the number of worker containers is
specified. Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The { \em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm. Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function will be tested. {\em Instances}  has a default value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of function evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

@@

As a conclusion, EvoSwarm is a framework designed and tested according
to industry standards which is able to bring many benefits to the
bioinspired algorithm community, by showing the possibility of
creating new, high performance, mixed algorithms, which can also be
easily deployed to the cloud, or locally to systems that simply need
to have Docker and Docker Compose installed.

It also opens new possibilities. From the design point of view, it
would be interesting to try and use automatic scaling, instead of
setting the number of workers by hand. This would need a certain
amount of redesign, the most important of which would be using
container orchestration systems such as Kubernetes or Docker
Swarm. Instead of setting a fixed number of workers, only the maximum
amount would need to be established, which could be done directly or
by setting an evaluation budget.
